Optimal Algorithms for Dynamic Balanced Graph Partitioning

### 2020-05-13 paper/disc20clustering.tex
improvements

	We proceed by joining the current pivot node to a~larger ground set in a similar way,
	as long as a feasible partition exists afterwards.
	Consider the isolated node $x_i$ currently collocated with $B$.
	 We issue a request between $x_i$ and some node $y_i \in C_i$,
	 where $C_i$ is the largest ground set
	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$,
	 only if  a feasible partition exists
	 that collocates all nodes in $\{x_i\} \cup C_i $.
	Any such feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
	Otherwise,
	no such feasible partition exists and
	we terminate the process.

	 Once the process terminates,
	 there are at most $\ell$ isolated nodes left. 
	 Assume for contradiction that this is not the case,
	 i.e., there are at least $\ell+1$ isolated nodes left.
	 Then there must be two isolated nodes $u$ and $v$,
	  with the same cluster of origin,
	 i.e., $I(\{u\}) = I(\{v\})$.
	 Consider the partition $P$
	  obtained from the initial partition after swapping $x_0$ and $y_0$
	 with $u$ and $v$ (respectively).
	 In this partition,
	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{u,v\})$.
	 Note that after the first request $\{x_0,y_0\}$,
	   we issue requests only between nodes that have the same cluster of origin and
	   all these nodes are collocated in $P$.
	 Therefore all ground sets constructed so far are collocated in  $P$,
	 i.e., $P$ is a feasible partition,
	 which contradicts our assumption.
	 Hence, there are at most $\ell$ isolated nodes
	 at the point of termination.

	The number of isolated nodes decrease after each step.
	Consider the last step, $i^*$, when exactly $\ell+1$ isolated nodes exist before  $x_{i^*}$ joins to another ground set.
	I.e., there are less than $\ell+1$ isolated nodes after the step $i^*$.
	Then, at the end of step $i^*-1$,
	there exist two isolated nodes $x^*$ and $y^*$
	with the same cluster of origin, i.e., $I(\{x^*\})=I(\{y^*\})$.
%
    In what follows,
    we bound costs incurred by \OPT and \ALG over the  sequence of requests
    $\sigma := \{ \{x_0,y_0\}, \dots, \{x_{i^*-1},y_{i^*-1}\}  \}$.

@@

	\ALG performs at least one swap each time a~ground set grows,
	i.e., each time it joins the isolated node $x_i$ (collocated with $B$), to  the ground set $C_i$ as specified before.
%	In other words,
%	some ground set grows by one node at each of these steps.
	Consider any ground set $C^*$ after serving $\sigma$.
	This ground set has grown exactly $|C^*|-1$ times until the end of $\sigma$. 
	Let $\mathcal{C}^*$ be the set of all ground sets after serving the last request in $\sigma$.
	Thus, $\mathcal{C}^*$ includes ground sets $B$,
	$\{x_0,y_0\}$,
	and $\ell+1$ singleton ground sets.
%	(including $x^*$ and possibly $y^*$).	
	Each time we join the pivot node to the largest ground set of same origin.
	Hence, every  other ground set in $\mathcal{C}^*$ has grown maximally.
	There are at most $\ell-1$ such ground sets in addition to the ones 
	we considered so far,
	one per cluster except the cluster containing $B$. 
	Therefore,  $|\mathcal{C}^*| \leq 1 + 1 + (\ell+1)  + (\ell-1) = 2\ell+2$.
	Note that among all non-singleton ground sets in $\mathcal{C}^*$,
	 only $B$ does not grow during the process.
	 Thus,
    the total number of times that 
	a ground set in $\mathcal{C}^*$  has grown is
	\begin{align*}
	\sum_{C^* \in \mathcal{C}^* } (|C^*|-1) - (k-1) &= 
	\sum_{C^* \in \mathcal{C}^* } |C^*| - 
		\sum_{C^* \in \mathcal{C}^* } 1 - (k-1) \\
	&\geq  k  \ell - (2\ell+2) - (k-1) =  (k-2)(\ell-1) - 3,
	\end{align*}
	which is the least number of swaps performed by \ALG.
	The competitive ratio is then
	$\ALG / \OPT \geq ((k-2)(\ell-1) - 3) / 2$.
	%%
	\begin{comment}
	\ALG performs at least one swap  every time   a~ground set is constructed.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
	\end{comment}
	%%
\end{proof}

>>>>

	We proceed by joining the current pivot node to a~larger ground set in a similar way,
	as long as a feasible partition exists afterward.
	Consider the isolated node $x_i$ currently collocated with $B$.
	 We issue a request between $x_i$ and some node $y_i \in C_i$,
	 where $C_i$ is the largest ground set
	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$,
	 only if  a feasible partition exists
	 that collocates all nodes in $\{x_i\} \cup C_i $.
	Any such feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
	Otherwise,
	no such feasible partition exists and
	we terminate the process.

	 Once the process terminates,
	 there are at most $\ell$ isolated nodes left. 
	 Assume for contradiction that this is not the case,
	 i.e., there are at least $\ell+1$ isolated nodes left.
	 Then there must be two isolated nodes $u$ and $v$,
	  with the same cluster of origin,
	 i.e., $I(\{u\}) = I(\{v\})$.
	 Consider the partition $P$
	  obtained from the initial partition after swapping $x_0$ and $y_0$
	 with $u$ and $v$ (respectively).
	 In this partition,
	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{u,v\})$.
	 Note that after the first request $\{x_0,y_0\}$,
	   we issue requests only between nodes that have the same cluster of origin and
	   all these nodes are collocated in $P$.
	 Therefore all ground sets constructed so far are collocated in  $P$,
	 i.e., $P$ is a feasible partition,
	 which contradicts our assumption.
	 Hence, there are at most $\ell$ isolated nodes after the termination.

	The number of isolated nodes decrease after each step.
	Consider the last step, $j$, when exactly $\ell+1$ isolated nodes exist before  $x_{j}$ joins to another ground set.
	That is, there are less than $\ell+1$ isolated nodes after the step $j$.
	Then, at the end of step $j-1$,
	there exist two isolated nodes $x^*$ and $y^*$
	with the same cluster of origin, i.e., $I(\{x^*\})=I(\{y^*\})$.
%
   Next,
    we bound costs incurred by \OPT and \ALG over the  sequence of requests
    $\sigma := \langle \{x_0,y_0\}, \dots, \{x_{j-1},y_{j-1}\}  \rangle$.

@@

	\ALG performs at least one swap each time a~ground set grows,
	i.e., each time it joins the isolated node $x_i$ (collocated with $B$) to  the ground set $C_i$ as specified before.
%	In other words,
%	some ground set grows by one node at each of these steps.
	Consider any ground set $C^*$ after serving $\sigma$.
	This ground set has grown exactly $|C^*|-1$ times until the end of $\sigma$. 
	Let $\mathcal{S}$ be the set of all ground sets after serving the last request in $\sigma$.
	Thus, $\mathcal{S}$ includes ground sets $B$,
	$\{x_0,y_0\}$,
	and $\ell+1$ singleton ground sets.
	Among the remaining ground sets in $\mathcal{S}$,
	no two ground sets have the same origin.
	Otherwise,
   the smaller ground set is either a singleton,
   which contradicts the bound $\ell+1$ on the number of singletons,
	or we have joined nodes to it at some step,
	contradicting our choice of the largest ground set of the same origin.
	Hence, every  other ground set in
	$\mathcal{S}$ has grown \emph{maximally} over $\sigma$.
	There are at most $\ell-1$ maximally grown ground sets,
	one per cluster,
	 excluding the cluster containing $B$. 
	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+1)  + (\ell-1) = 2\ell+2$.
	Note that among all non-singleton ground sets in $\mathcal{S}$,
	 only $B$ does not grow during the process.
	 Thus,
    the total number of times that 
	a ground set in $\mathcal{S}$  has grown is
	\begin{align*}
	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) &= 
	\sum_{C^* \in \mathcal{S} } |C^*| - 
		\sum_{C^* \in \mathcal{S} } 1 - (k-1) \\
	&\geq  k  \ell - (2\ell+2) - (k-1) =  (k-2)(\ell-1) - 3,
	\end{align*}
	which bounds the number of swaps performed by \ALG.
	The competitive ratio is then
	$\ALG / \OPT \geq ((k-2)(\ell-1) - 3) / 2$.
	%%
	\begin{comment}
	\ALG performs at least one swap  every time   a~ground set is constructed.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
	\end{comment}
	%%
\end{proof}
